# Abundance

Here is a review of existing methods.

## Lincoln-Peterson

## Full Likelihood

## Conditional Likelihood

The conditional likelihood abundance estimator proposed by @huggins_statistical_1989 and @alho_logistic_1990, which was further extended in @huggins_practical_1991, is an extension to previous abundance estimators to account for heterogeneous capture probabilities ($p$). The model estimates individual capture probabilities and abundance conditional on captured individuals.


### Algebra
$$  \mathcal{L} = \prod_{i=1}^n \prod_{t=1}^t p_{i,t}^{z_{i,t}}(1-p_{i,t})^{1-z_{i,t}} $$
where, \linebreak
  $z_{i,t}$ takes the value of 1 if individual $i$ has been captured before occasion $t$ and 0 otherwise.

Thus the capture history $y_{i,t}$ is used to estimate the capture probability of individual $i$ as a Bernoulli trial,
$$y_{i,t} \sim Bernoulli(p_{i,t}) $$
and abundance $\hat{N}$ is derived condtional on the count of known individuals ($C$), sometimes referred to the minimum number of known alive ($MNKA$).
$$\hat{N} = \frac{C}{1-\prod^{t}(1-p_t)}$$
Variation in detection probability can be modeled using linear logistic models or other variations used to estimate probabilities 0-1.

### Simulation 

```{r simulation}
N <- 150 #True population size
n_occ <- 4 #Number of trapping occasions
p <- 0.50 #Probability of first detection

true_detections <- array(NA, dim=c(N,n_occ))
for (t in 1:n_occ){
  true_detections[,t] <- rbinom(n=N,size=1,prob=p)
}
observed <- true_detections[apply(true_detections,1,max) == 1,]
MNKA <- nrow(observed)
print( paste0("Number ever detected: ", MNKA,sep = " ") ) #number ever detected
```
### Models {.tabset}

::: {.tab}
<button class="tablinks" onclick="unrolltab(event,'JAGS')">JAGS</button>
<button class="tablinks" onclick="unrolltab(event, 'Stan')">Stan</button>
<button class="tablinks" onclick="unrolltab(event, 'Greta')">Greta</button>
::: {#JAGS .tabcontent}
#### JAGS model fit {-}
```{r,warning=FALSE,message=FALSE,error=FALSE, results='hide',cache = TRUE}
library(knitr)
library(R2jags)
data <- list(
  y=observed,
  nsites=nrow(observed),
  MNKA=MNKA,
  n_occ=n_occ
)
model_string <- textConnection(
  "
  model {
  # Likelihood
  for(i in 1:nsites) {
    # Observation model
    for(j in 1:4) {
      y[i, j] ~ dbern(p)
    }
  }
  for(t in 1:4){
    p_un[t] <- (1-p)
  }
  # Priors
  p ~ dunif(0, 1) # Uninformative prior
  # Derived values
  N <- (MNKA / (1-prod(p_un[])))
}")
parameters <- c("p","N")
inits <- function() {
  list( 
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, inits, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)
jags_output <- cbind(Sampler=rep('jags',2), round(model$BUGSoutput$summary[c(1,3),c(1,2,3,7)],2))
```
:::
::: {#Stan .tabcontent}
#### Stan  {-}
```{r ,warning=FALSE,message=FALSE,error=FALSE, results='hide',cache = TRUE}
library(knitr)
library(rstan)
data <- list(
  y=observed,
  nsites=nrow(observed),
  MNKA=MNKA,
  n_occ=n_occ
)
stan_model <- "
data {
  int<lower=0> MNKA;
  int<lower=0> nsites;
  int<lower=0> n_occ;
  int<lower=0,upper=1> y[MNKA, n_occ];
}
parameters {
  real<lower=0, upper=1> p;
}
model {  
 for(i in 1:nsites)
    for(j in 1:4)
      y[i, j] ~ bernoulli(p);
}
generated quantities {
  real N = MNKA / (1-(1-p)^4);
}
"
nc <- 4
stan.samples <- stan(model_code = stan_model, data = data, iter = 10000, chains = nc, cores = nc, open_progress=FALSE)
stan_output <- cbind(Sampler=rep('stan',2,), round(summary(stan.samples)$summary[c(2,1),c(1,3,4,8)],2))
```
:::
::: {#Greta .tabcontent}
#### Greta  {-}
```{r,warning=FALSE,message=FALSE,error=FALSE, results='hide',cache = TRUE}
library(greta)
#capture_vec <- unlist(observed)
p_greta <- beta(1, 1)
# likelihood
capture_vec <- as_data(observed)
distribution(capture_vec) <- bernoulli(p_greta)
# derived parameter
pstar <-  1-(1-p_greta)*(1-p_greta)*(1-p_greta)*(1-p_greta)
N_hat <-  MNKA / pstar
# defining the model
m <- model(p_greta, N_hat,pstar)

# sampling
draws <- greta::mcmc(m, n_samples = 1000)
greta_output <- cbind(Sampler=rep('greta',2),round(summary(draws)$statistics[c(2,1),c(1,2)],2),round(summary(draws)$quantiles[c(2,1),c(1,5)],2))
```
```{r}
plot(m)
head(observed)
```
:::
:::

### Comparison
```{r Comparison}
results <- rbind(jags_output,
                 stan_output,
                 greta_output)
kable(results)
```

## Data Augmentation
## N-Mixture Model
## Spatial Capture-Recapture

    