# Abundance



## Lincoln-Peterson

Lincoln 1930, Petersen 1896

The Lincoln-Petersen abundance estimator arises in cases where we have mark-recapture data recorded over two capture occasions. During the first occasion individuals are captured and uniquely marked. During the second occasion the number of individuals captured is recorded, along with the number of individuals marked during the first occasion that are recaptured. 

### Algebra

The data consist of:
- $M_1$ : the number of individuals captured and uniquely marked during the first occasion
- $M_2$ : the number of previously marked individuals captured during the second occasion
- $C$ : the number of animals captured during the second trapping occasion

$$M_2 \sim Binomial(C, \frac{M_1}{\hat{N}}) $$
$$\mathcal{L}(C| M_1, p = \frac{M_1}{\hat{N}}) = {M_1 \choose C} p^C(1-p)^{M_1-C}$$
The maximum likelihood estimate for $\hat{N}$ is, 
$$\hat{N} = \frac{M_1*C}{M_2}$$

### Simulation 
```{r}
set.seed(1)
N <- 75 #True population size
n_occ <- 2 #Number of trapping occasions
p <- 0.50 #Probability of first detection

true_detections <- array(NA, dim=c(N,n_occ))
for (t in 1:n_occ){
  true_detections[,t] <- rbinom(n=N,size=1,prob=p)
}
observed <- true_detections[apply(true_detections,1,max) == 1,]
M_1 <- sum(observed[,1])
M_2 <- nrow(subset(observed,rowSums(observed)>1))
C <- sum(observed[,2])
```
### Models {.tabset}

::: {.tab}
<button class="tablinks" onclick="unrolltab(event,'JAGS')">JAGS</button>
<button class="tablinks" onclick="unrolltab(event,'NIMBLE')">NIMBLE</button>
<button class="tablinks" onclick="unrolltab(event, 'Stan')">Stan</button>
<button class="tablinks" onclick="unrolltab(event, 'Greta')">Greta</button>
::: {#JAGS .tabcontent}
#### JAGS model fit {-}
```{r ,eval = FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(knitr)
library(R2jags)
data <- list(
          M_1 = M_1,
          M_2= M_2,
          C = C,
          LB = C- M_1 + M_2,
          UB = 5*C
)
model_string <- textConnection(
  "
  model {
  
  # Likelihood
  M_2 ~ dbin(M_1/N, C )
  # Priors
  N ~ dunif(LB, UB) # Slightly informative prior

  # Derived values
  }
"
)
parameters <- c("N")
set_initial_value <- function() {
  list( 
    
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, set_initial_value, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)
```
:::
::: {#NIMBLE .tabcontent}
#### NIMBLE model fit {-}
```{r ,eval = FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(nimble)
n_data <- list(
          M_1 = M_1,
          M_2= M_2,
          C = C
)
n_constants <- list(
          LB = C- M_1 + M_2,
          UB = 5*C
)
Nimble_Code <- nimbleCode({
  # Likelihood
  M_2 ~ dbin(M_1/N, C )
  # Priors
  N ~ dunif(LB, UB) # Slightly informative prior

  # Derived values

})

n_params <- c("N")
n_inits <- list( 
  
)
Nimble_Model <- nimbleModel(
  code = Nimble_Code,
  constants = n_constants,
  data = n_data,
  inits = n_inits
)
MCMC_Model <- configureMCMC(Nimble_Model, monitors = n_params, print = T, enableWAIC = F)
Model1_MCMC <- buildMCMC(MCMC_Model)
Comp_Model <- compileNimble( Nimble_Model, showCompilerOutput = TRUE )
Comp_Model <- compileNimble(Model1_MCMC, project = Nimble_Model)

niter=10000
Model_samples <- runMCMC(Comp_Model, niter = niter, nburnin=niter/2,nchains=3,summary = TRUE)
# mcmc_combo(Model_samples$samples, pars = c("N", "p"))
# round(Model_samples$summary$all.chains,2)

```
:::
::: {#Stan .tabcontent}
#### Stan  {-}
```{r ,eval = FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(knitr)
library(rstan)

data <- list(
          M_1 = M_1,
          M_2= M_2,
          C = C
)

stan_model <- "
data {
  int<lower=0> M_1;
  int<lower=0> M_2;
  int<lower=0> C;
}
parameters {
  real<lower=(C - M_2 + M_1)> N;
}
model {
  M_2 ~ binomial(C, M_1 / N);
}
"
nc <- 4

stan.samples <- stan(model_code = stan_model, data = data, iter = 10000, chains = nc, cores = nc)
```

:::
::: {#Greta .tabcontent}
#### Greta  {-}
```{r ,eval = FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(reticulate)
reticulate::use_condaenv("r-reticulate")
reticulate::py_config()
library(greta)
LB <- C- M_1 + M_2
UB <- 5*C
# priors
N <- uniform(LB,UB)
# likelihood
distribution(M_2) <- binomial(C, M_1 / N)
# derived parameter

# defining the model
m <- model(N) #objects to sample
# sampling
draws <- greta::mcmc(m, n_samples = 1000)
```
```{r,eval = FALSE}
plot(m)
```
:::
:::

### Comparison


## Full Likelihood

## Conditional Likelihood

The conditional likelihood abundance estimator proposed by @huggins_statistical_1989 and @alho_logistic_1990, which was further extended in @huggins_practical_1991, is an extension to previous abundance estimators to account for heterogeneous capture probabilities ($p$). The model estimates individual capture probabilities and abundance conditional on captured individuals.


### Algebra
The capture history $y_{i,t}$ is used to estimate the capture probability of individual $i$ as a Bernoulli trial,
$$y_{i,t} \sim Bernoulli(p_{i,t}) $$
$$\mathcal{L}(p| y) = \prod_{i=1}^n \prod_{t=1}^t p_{i,t}^{z_{i,t}}(1-p_{i,t})^{1-z_{i,t}}$$

Abundance $\hat{N}$ is derived conditional on the count of known individuals ($C$), sometimes referred to the minimum number of known alive ($MNKA$).
$$\hat{N} = \frac{C}{1-\prod^{t}(1-p_t)}$$
Variation in detection probability can be modeled using linear logistic models or other variations used to estimate probabilities 0-1.

### Simulation 

```{r}
set.seed(1)
N <- 150 #True population size
n_occ <- 4 #Number of trapping occasions
p <- 0.50 #Probability of first detection

true_detections <- array(NA, dim=c(N,n_occ))
for (t in 1:n_occ){
  true_detections[,t] <- rbinom(n=N,size=1,prob=p)
}
observed <- true_detections[apply(true_detections,1,max) == 1,]
MNKA <- nrow(observed)
print( paste0("Number ever detected: ", MNKA,sep = " ") ) #number ever detected
```
### Models {.tabset}

::: {.tab}
<button class="tablinks" onclick="unrolltab(event,'JAGS')">JAGS</button>
<button class="tablinks" onclick="unrolltab(event,'NIMBLE')">NIMBLE</button>
<button class="tablinks" onclick="unrolltab(event, 'Stan')">Stan</button>
<button class="tablinks" onclick="unrolltab(event, 'Greta')">Greta</button>
::: {#JAGS .tabcontent}
#### JAGS model fit {-}
```{r ,eval = FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(R2jags)
data <- list(
  y=observed,
  n_sites=nrow(observed),
  MNKA=MNKA,
  n_occ=n_occ
)
model_string <- textConnection(
  "
  model {
  # Likelihood
  for(i in 1:n_sites) {
    # Observation model
    for(j in 1:n_occ) {
      y[i, j] ~ dbern(p)
    }
  }
  for(t in 1:n_occ){
    p_un[t] <- (1-p)
  }
  # Priors
  p ~ dunif(0, 1) # Uninformative prior
  # Derived values
  N <- (MNKA / (1-prod(p_un[])))
}")
parameters <- c("p","N")
inits <- function() {
  list( 
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, inits, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)
```
:::
::: {#NIMBLE .tabcontent}
#### NIMBLE model fit {-}
```{r ,eval = FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(nimble)
n_data <- list(
  y = observed,
  MNKA = MNKA
)
n_constants <- list(
  n_occ = n_occ
)
Nimble_Code <- nimbleCode({
  # Likelihood
  for(i in 1:MNKA) {
    # Observation model
    for(j in 1:n_occ) {
      y[i, j] ~ dbern(p)
    } #j
  } #i
  
  # Priors
  p ~ dunif(0, 1) # Uninformative prior
  
  # Derived values
  for(t in 1:n_occ){
    p_un[t] <- (1-p)
  } #t
  N <- (MNKA / (1-prod(p_un[1:n_occ]))) #The only difference in this model is here declaring dimensions
})

n_params <- c("p", "N")
n_inits <- list( 
  
)
Nimble_Model <- nimbleModel(
  code = Nimble_Code,
  constants = n_constants,
  data = n_data,
  inits = n_inits
)
MCMC_Model <- configureMCMC(Nimble_Model, monitors = n_params, print = T, enableWAIC = F)
Model1_MCMC <- buildMCMC(MCMC_Model)
Comp_Model <- compileNimble( Nimble_Model, showCompilerOutput = TRUE )
Comp_Model <- compileNimble(Model1_MCMC, project = Nimble_Model)

niter=10000
Model_samples <- runMCMC(Comp_Model, niter = niter, nburnin=niter/2,nchains=3,summary = TRUE)
#mcmc_combo(Model_samples$samples, pars = c("N", "p"))
round(Model_samples$summary$all.chains,2)
```
:::
::: {#Stan .tabcontent}
#### Stan model fit {-}
```{r ,eval = FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(knitr)
library(rstan)
data <- list(
  y=observed,
  nsites=nrow(observed),
  MNKA=MNKA,
  n_occ=n_occ
)
stan_model <- "
data {
  int<lower=0> MNKA;
  int<lower=0> nsites;
  int<lower=0> n_occ;
  int<lower=0,upper=1> y[MNKA, n_occ];
}
parameters {
  real<lower=0, upper=1> p;
}
model {  
 for(i in 1:nsites)
    for(j in 1:4)
      target += bernoulli_lpmf(y[i, j] | p);
}
generated quantities {
  real pstar = (1-(1-p)^n_occ);
  real N = MNKA / pstar;
}
"
nc <- 4
stan.samples <- stan(model_code = stan_model, data = data, iter = 10000, chains = nc, cores = nc, open_progress=FALSE)
```
:::
::: {#Greta .tabcontent}
#### Greta model fit{-}
```{r ,eval = FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(reticulate)
reticulate::use_condaenv("r-reticulate")
reticulate::py_config()
library(greta)
capture_vec <- unlist(observed)
# priors
p_greta <- beta(1, 1)
# likelihood
distribution(capture_vec) <- bernoulli(p_greta)
# derived parameters
pstar <- 1 - (1 - p_greta)^n_occ
N_hat <- MNKA / pstar
# defining the model
m <- model(p_greta, N_hat, pstar)
# sampling
draws <- greta::mcmc(m, n_samples = 1000)
```
```{r,eval = FALSE}
plot(m)
```
:::
:::

### Comparison
```{r, echo=FALSE,warning=FALSE}
# round(model$BUGSoutput$summary[c(1,3),c(1,2,3,7)],2) #jags
# round(summary(stan.samples)$summary[c(2,1),c(1,3,4,8)],2) #stan
# summary(draws) #greta
library(gt)
library(tidyverse)
results_df <- data.frame(
              Sampler=c("Simulated","JAGS","NIMBLE","Stan","greta","Simulated","JAGS","NIMBLE","Stan","greta"),
              Parameter=c("N","N","N","N","N","p","p","p","p","p"),
              Mean=c(150,149.73,149.73,149.73,149.73,NA,0.52,0.52,0.52,0.52),
              SD=c(NA,1.46,1.46,1.43,1.44,NA,0.02,0.02,0.02,0.02),
              LCL=c(NA,147.21,147.21,147.25,147.22,NA,0.48,0.48,0.48,0.48),
              UCL=c(NA,152.95,152.95,152.79,152.96,NA,0.57,0.57,0.57,0.57)
              )

# Create a gt table based on preprocessed
# `sp500` table data
results_df %>%
  dplyr::select(-Sampler,-Parameter) %>%
  gt() %>%
  tab_row_group(
    group = "p",
    rows = 5:8
  ) %>%
  tab_row_group(
    group = "N",
    rows = 1:4
  ) %>%
  tab_header(
    title = "Estimate Comparison",
    subtitle = "Comparison of Huggins abundance (N) and detection 
              \n probability (p) estimates fit using JAGS, Stan, greta."
  )
```


## Data Augmentation
## N-Mixture Model
## Distance Sampling
## Spatial Capture-Recapture
## Time to Event
    
