# (APPENDIX) Appendix {-} 

# Appendix A
## Authors Guidelines
We have decided to adapt the tidyverse style guide. Detailed content can be found here.  
  
[tidyverse style guide][1]
  
Before committing code to the git repository, it should be styled using the styler package.

Primary points are outlined below:
- Use underscores to separate words in both variable and file names. 
````
File Names:
chapter_1.Rmd
---
Variable Names:
observed_data
````
- When submitting a description of a new model use the template provided in Appendix B. However, if extenuating circumstances make the template in Appendix B untenable for a model, describe the reason for departure in the git commit message.
- Variables should be named with nouns, functions should be named with verbs
- Use dots to separate words in function names
````
simulate.code <- function(X) {
}
````
- Label function arguments
- Spaces after commas not before
- Do not pad parentheses with spaces
- Pad operators with space
````
foo == bar
foo <- bar
foo + bar
foo * bar
````
[1]: https://style.tidyverse.org/

# Appendix B
## Model Template

````
### Model Name 
Model description

### Algebra

Algebraic model description in latex

### Simulation 
```{r simulation}`r ''`

simulated_data <- 1

```  

### JAGS
```{r JAGS,warning=FALSE,message=FALSE,error=FALSE, results='hide',cache = TRUE}`r ''`
library(R2jags)
data <- list(

)
model_string <- textConnection(
  "
  model {
  
  # Likelihood
 
  # Priors


  # Derived values

  }
"
)
parameters <- c( )
set_initial_value <- function() {
  list( 
    
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, set_initial_value, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)
```

```{r output_jags_table}`r ''`
library(knitr)
kable(round( ,2))
```

### Stan
```{r stan_model_fit ,warning=FALSE,message=FALSE,error=FALSE, results='hide',cache = TRUE}`r ''`
library(rstan)

data <- list(

)

stan_model <- "
data {

}

parameters {

}

model {  

}

generated quantities {

}
"
nc <- 4

stan.samples <- stan(model_code = stan_model, data = data, iter = 10000, chains = nc, cores = nc)
```

```{r output_stan_table}`r ''`
library(knitr)
kable(round( ,2))
```

### Comparison
```{r Comparison}`r ''`

```
````