# (APPENDIX) Appendix {-} 

# Appendix A
## Authors Guidelines
We have decided to adapt the tidyverse style guide. Detailed content can be found here.  
  
[tidyverse style guide][1]
  
Before committing code to the git repository, it should be styled using the styler package.

Primary points are outlined below:
- Use underscores (snake_case) to separate words in both variable and file names. 
- Use periods to separate words in function names
````
File Names:
chapter_1.Rmd
---
Variable Names:
observed_data
````
- When submitting a description of a new model use the template provided in Appendix B. However, if extenuating circumstances make the template in Appendix B untenable for a model, describe the reason for departure in the git commit message.
- Variables should be named with nouns, functions should be named with verbs
- Use dots to separate words in function names
````
simulate.code <- function(X) {
}
````
- Label function arguments
- Spaces after commas not before
- Do not pad parentheses with spaces
- Pad operators with space
````
foo == bar
foo <- bar
foo + bar
foo * bar
````
[1]: https://style.tidyverse.org/

# Appendix B
## Model Template

````
## Model Name 
Model description

### Algebra

Algebraic model description in latex

### Simulation {.tabset} 

::: {.tab}
<button class="tablinks" onclick="unrolltab(event,'R_XXX_sim')">R</button>
<button class="tablinks" onclick="unrolltab(event,'Python_XXX_sim')">Python</button>
<button class="tablinks" onclick="unrolltab(event,'Julia_XXX_sim')">Julia</button>
::: {#R_XXX_sim .tabcontent}
#### {-}
<br>
```{r , eval = FALSE}r ''`
```
:::
::: {#Python_XXX_sim .tabcontent}
#### {-}
<br>
```{python , eval = FALSE}r ''`

```
:::
::: {#Julia_XXX_sim .tabcontent}
#### {-}
<br>
```{julia , eval = FALSE}r ''`

```
:::
:::

### Models {.tabset}

::: {.tab}
<button class="tablinks" onclick="unrolltab(event,'JAGS_XXX_mod')">JAGS</button>
<button class="tablinks" onclick="unrolltab(event,'NIMBLE_XXX_mod')">NIMBLE</button>
<button class="tablinks" onclick="unrolltab(event, 'Stan_XXX_mod')">Stan</button>
<button class="tablinks" onclick="unrolltab(event, 'Greta_XXX_mod')">Greta</button>
<button class="tablinks" onclick="unrolltab(event,'Pymc3_XXX_mod')">Pymc3</button>
<button class="tablinks" onclick="unrolltab(event,'Turing_XXX_mod')">Pymc3</button>
::: {#JAGS_XXX_mod .tabcontent}
#### {-}
<br>
```{r ,eval = FALSE}r ''`
library(knitr)
library(R2jags)
data <- list(

)
model_string <- textConnection(
  "
  model {
  
  # Likelihood
 
  # Priors


  # Derived values

  }
"
)
parameters <- c( )
set_initial_value <- function() {
  list( 
    
  )
}
ni <- 10000 ; nt <- 1 ; nc <- 3
model <- jags(data = data, 
              inits = inits, 
              parameters = parameters, 
              model.file = model_string, 
              n.chains = nc, 
              n.thin = nt, 
              n.iter = ni, 
              n.burnin = ni/2)
```
:::
::: {#NIMBLE_XXX_mod .tabcontent}
#### {-}
<br>
```{r ,eval = FALSE}r ''`
library(nimble)
n_data <- list(

)
n_constants <- list(

)
Nimble_Code <- nimbleCode({
  # Likelihood
  
  # Priors
  
  # Derived values

})

n_params <- c( )
n_inits <- list( 
  
)
Nimble_Model <- nimbleModel(
  code = Nimble_Code,
  constants = n_constants,
  data = n_data,
  inits = n_inits
)
MCMC_Model <- configureMCMC(Nimble_Model, monitors = n_params, print = T, enableWAIC = F)
Model1_MCMC <- buildMCMC(MCMC_Model)
Comp_Model <- compileNimble( Nimble_Model, showCompilerOutput = TRUE )
Comp_Model <- compileNimble(Model1_MCMC, project = Nimble_Model)

ni <- 10000 ; nt <- 1 ; nc <- 3
Model_samples <- runMCMC(mcmc = Comp_Model, 
                         niter = ni, 
                         nburnin = ni/2,
                         nchains = nc,
                         summary = TRUE)
mcmc_combo(Model_samples$samples, pars = c("N", "p"))
round(Model_samples$summary$all.chains,2)

```
:::
::: {#Stan_XXX_mod .tabcontent}
#### {-}
<br>
```{r ,eval = FALSE}r ''`
library(knitr)
library(rstan)

data <- list(

)

stan_model <- "
data {

}

parameters {

}

model {  

}

generated quantities {

}
"
ni <- 10000 ; nt <- 1 ; nc <- 4
stan.samples <- stan(model_code = stan_model, 
                     data = data, 
                     iter = ni, 
                     warmup = floor(ni/2),
                     chains = nc, 
                     cores = nc)
```

:::
::: {#Greta_XXX_mod .tabcontent}
#### {-}
<br>
```{r ,eval = FALSE}r ''`
library(greta)
# priors

# likelihood

# derived parameter

# defining the model
m <- model() #objects to sample

# sampling
ni <- 10000 ; nt <- 1 ; nc <- 4
draws <- greta::mcmc(model = m, 
                     n_samples = ni,
                     thin = nt,
                     warmup = floor(ni/2),
                     chains = nc, 
                     n_cores = nc)
```
:::
::: {#Pymc3_XXX_mod .tabcontent}
#### {-}
<br>
```{python ,eval = FALSE}r ''`
with Model() as model: 
    # Priors for unknown model parameters
    
    # Likelihood (sampling distribution) of observations
   
    trace = pm.sample()

```
:::
::: {#Turing_XXX_mod .tabcontent}
#### {-}
<br>
```{julia ,eval = FALSE}r ''`
using Turing, Distributions
@model function linear_regression(x, y)

end

model = linear_regression(x, y)
chain = sample(model, NUTS(0.65), 3_000);
```
:::
:::

### Comparison
```{r , echo=FALSE, warning=FALSE}r```
#round(model$BUGSoutput$summary[c(1,3),c(1,2,3,7)],2) #jags 
#round(summary(stan.samples)$summary[c(2,1),c(1,3,4,8)],2) #stan 
#summary(draws) #greta 
#https://gt.rstudio.com/articles/intro-creating-gt-tables.html

library(gt)
library(tidyverse)
results_df <- data.frame(
              Sampler=c(),
              Parameter=c(),
              Mean=c(),
              SD=c(),
              LCL=c(),
              UCL=c()
              )

results_df %>%
  dplyr::select(-Sampler,-Parameter) %>%
  gt() %>%
  tab_row_group(
    group = "p",
    rows = 5:8
  ) %>%
  tab_row_group(
    group = "N",
    rows = 1:4
  ) %>%
  tab_header(
    title = "",
    subtitle = ""
  )
```
````