--- 
title: "Quantitative Ecology"
subtitle: "Modelling applications in population ecology"
author: "Matt Weldy"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Setup

## Setup Python

```{r python, include=FALSE}
library(reticulate)
#Sys.setenv(RETICULATE_PYTHON = "/home/mweldy/anaconda3/envs/r-QEW/bin/python")
# conda create -n r-QEW python=3.7 tensorflow=1.14 pyyaml requests Pillow pip numpy=1.16 h5py=2.8 tensorflow-probability=0.7 PyMC3=3.11.2 arviz 
# conda install -c conda-forge pymc3=3.11.2
reticulate::py_config()

# %config InlineBackend.figure_format = 'retina'
# az.style.use("arviz-darkgrid")
# 
# file.edit(file.path("~", ".Rprofile")) # edit .Rprofile in HOME
# file.edit(".Rprofile") # edit project specific .Rprofil

# 
# plt.figure(figsize=(7, 7))
# traceplot(trace)
# plt.tight_layout();
```
## Setup Julia
```{r setup_julia, include=FALSE}
library(JuliaCall)
knitr::opts_chunk$set(echo = TRUE)
julia_setup(JULIA_HOME = "/opt/julia-1.6.0/bin")
julia <- julia_setup()
```

Was the setup correct
```{julia}
a = sqrt(2)
```

# Prerequisites

This is a work in progress. 

The scope of this work is to demonstrate the fit of well known ecological models using a variety of tools. Our aim is to provide brief model descriptions, primary citations, and a simple simulation and model fit.

Contributions to this web page are welcome, and can be made through the books github repository. The website is hosted through netlify, which offers continuous integration with the books git repository. Commit changes to the repository after using the bookdown and changes will populate to the page. Please try to structure indivudal model fits using the template outlined in Appendix B.

```{r , eval=FALSE}
bookdown::render_book("index.Rmd", "bookdown::gitbook")
```

```{r, include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
