# Introduction {#intro}

## Probabilistic Modelling in Ecology

## Common Distributions

## Uncommon Distributions

## Linear Models

```{r}
library(lme4)
library(R2jags)
library(runjags)
#library(equatiomatic)
```


### Model of the Mean 

This is an intercept only model. We are estimating two parameters, $\mu$ and $\sigma$. 

Average length of adult coho salmon 75 cm with standard deviation of 20.

```{r}
n <- 1000
mu <- 75
sd <- 20
y <- rnorm(n, mean = mu, sd = sd)
```

```{r}
hist(y)
```

$$y \sim Normal(\mu, \sigma)$$
$$mu = \beta_0$$
$$P(\mu, \sigma| y) \propto \mathbb(L)*P$$

```{r , eval = FALSE}
lm_fit <- lm(y ~ 1)
data <- list(
            n = n,
            y = y
)
model_string <- textConnection(
  "
  model {
  # Likelihood
  for (i in 1:n){
    y[i] ~ dnorm(mu[i], tau)
    residuals[i] <- y[i] - mu[i]
    mu[i] <- intercept
  }
  # Priors
  intercept ~ dnorm(0, 0.00001)
  sigma ~ dunif(0, 100) # standard deviation
	tau <- 1 / (sigma * sigma) # sigma^2 doesn't work in JAGS
	
  # Derived values
  residual_sum_sq <- sum(residuals[]^2)
  }
"
)
parameters <- c("intercept", "sigma","residual_sum_sq")
set_initial_value <- function() {
  list( 
    
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, set_initial_value, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)

```

### Difference of Means

This is a one-way anova with the intercept as the identity and the beta estimates representing the offset effects from the intercept. 

```{r}
n <- 1000
beta_0 <- 1.5
beta_1 <- 3
x <- c(rep(0,n/2),rep(1,n/2))
mu <- beta_0 + beta_1*x
sd <- 10
y <- rnorm(n, mean = mu, sd = sd)
#model.matrix(~x)
```

```{r , eval = FALSE}
lm_fit <- lm(y ~ x)
data <- list(
            n = n,
            y = y,
            x = x
)
model_string <- textConnection(
  "
  model {
  
  # Likelihood
  for (i in 1:n){
    y[i] ~ dnorm(mu[i], tau)
    residuals[i] <- y[i] - mu[i]
    mu[i] <- intercept + beta_1 * x[i]
  }
  # Priors
  intercept ~ dnorm(0, 0.00001)
  beta_1 ~ dnorm(0, 0.00001)
  sigma ~ dunif(0, 100) # standard deviation
	tau <- 1 / (sigma * sigma) # sigma^2 doesn't work in JAGS
  

  # Derived values
  residual_sum_sq <- sum(residuals[]^2)
  }
"
)
parameters <- c("intercept", "beta_1", "sigma", "residual_sum_sq" )
set_initial_value <- function() {
  list( 
    
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, set_initial_value, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)
```

### One-way ANOVA

This is a one-way anova with the intercept as the identity and the beta estimates representing the offset effects from the intercept. 

```{r}
n <- 1000
beta_0 <- 1.5
beta_1 <- 3
beta_2 <- -2
beta_3 <- 0.2
x <- as.factor(c(rep(1,n/4),rep(2,n/4),rep(3,n/4),rep(4,n/4)))
X <- model.matrix(~x)
head(X)
mu <- beta_0 + beta_1*X[,2] + beta_2*X[,3] + beta_3*X[,4]
sd <- 10
y <- rnorm(n, mean = mu, sd = sd)
```

```{r , eval = FALSE}
lm_fit <- lm(y ~ x)
data <- list(
            n = n,
            y = y,
            x = x
)
model_string <- textConnection(
  "
  model {
  # Likelihood
  for (i in 1:n){
    y[i] ~ dnorm(mu[i], tau)
    residuals[i] <- y[i] - mu[i]
    mu[i] <- intercept + betas[x[i]]
  }
  # Priors
  intercept ~ dnorm(0, 0.00001)
  betas[1] <- 0
  for(i in 2:4) {
    betas[i] ~ dnorm(0, 0.00001)
  }
  sigma ~ dunif(0, 100) # standard deviation
	tau <- 1 / (sigma * sigma) # sigma^2 doesn't work in JAGS
  # Derived values
  residual_sum_sq <- sum(residuals[]^2)
  }
"
)
parameters <- c("intercept", "betas", "sigma", "residual_sum_sq")
set_initial_value <- function() {
  list( 
    
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, set_initial_value, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)
```

### One-way ANOVA Identity

This is a one-way anova with an identity parameterization. Beta estimates represent the mean effect for each factor level. 

```{r , eval = FALSE}
n <- 1000
beta_0 <- 3
beta_1 <- 6
beta_2 <- -4
beta_3 <- 0
x <- as.factor(c(rep(1,n/4),rep(2,n/4),rep(3,n/4),rep(4,n/4)))
X <- model.matrix(~x-1)
tail(X)
mu <- beta_0*X[,1] + beta_1*X[,2] + beta_2*X[,3] + beta_3*X[,4]
sd <- 10
y <- rnorm(n, mean = mu, sd = sd)
```

```{r}
lm_fit <- lm(y ~ 0 + x)
data <- list(
            n = n,
            y = y,
            x = x
)
model_string <- textConnection(
  "
  model {
  # Likelihood
  for (i in 1:n){
    y[i] ~ dnorm(mu[i], tau)
    residuals[i] <- y[i] - mu[i]
    mu[i] <- betas[x[i]]
  }
  # Priors
  for(i in 1:4) {
    betas[i] ~ dnorm(0, 0.00001)
  }
  sigma ~ dunif(0, 100) # standard deviation
	tau <- 1 / (sigma * sigma) # sigma^2 doesn't work in JAGS
  # Derived values
  residual_sum_sq <- sum(residuals[]^2)
  }
"
)
parameters <- c("betas","sigma","residual_sum_sq")
set_initial_value <- function() {
  list( 
    
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, set_initial_value, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)
```

### Two-way ANOVA
```{r}
n <- 1000
beta_0 <- 3
beta_1 <- 6
beta_2 <- -2
beta_3 <- 3
x <- as.factor(c(rep(1,n/2),rep(2,n/2)))
x2 <- as.factor(c(rep(1,n/4),rep(2,n/4),rep(1,n/4),rep(2,n/4)))
X <- model.matrix(~ x + x2 + x:x2)
dim(X)
unique(X)
mu <- beta_0*X[,1] + beta_1*X[,2] + beta_2*X[,3] + beta_3*X[,4]
sd <- 3
y <- rnorm(n, mean = mu, sd = sd)
```

```{r , eval = FALSE}
lm_fit <- lm(y ~ x + x2 + x:x2)
data <- list(
            n = n,
            y = y,
            x = x,
            x2 = x2
)
model_string <- textConnection(
  "
  model {
  # Likelihood
  for (i in 1:n){
    y[i] ~ dnorm(mu[i], tau)
    residuals[i] <- y[i] - mu[i]
    mu[i] <- intercept + betas[x[i],x2[i]]
  }
  # Priors
  intercept ~ dnorm(0, 0.00001)
  betas[1,1] <- 0
  betas[1,2] ~ dnorm(0, 0.00001)
  betas[2,1] ~ dnorm(0, 0.00001)
  betas[2,2] ~ dnorm(0, 0.00001)
  
  sigma ~ dunif(0, 100) # standard deviation
	tau <- 1 / (sigma * sigma) # sigma^2 doesn't work in JAGS
  # Derived values
  residual_sum_sq <- sum(residuals[]^2)
  }
"
)
parameters <- c("intercept","betas","sigma","residual_sum_sq")
set_initial_value <- function() {
  list( 
    
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, set_initial_value, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)
```
### ANCOVA
```{r}
n <- 1000
beta_0 <- 2
beta_1 <- -2
beta_2 <- 3
beta_3 <- -1
x <- as.factor(c(rep(1,n/2),rep(2,n/2)))
year <- round(runif(n,1,10)) 
X <- model.matrix(~ x + year + x:year)
tail(X)
unique(X)
mu <- beta_0*X[,1] + beta_1*X[,2] + beta_2*X[,3] + beta_3*X[,4]
sd <- 20
y <- rnorm(n, mean = mu, sd = sd)
#df <- data.frame(y=y,x=x,year=year)
#template.jags(y~ x + year + x:year, data = df)
```

```{r , eval = FALSE}
lm_fit <- lm(y ~ x + year + x:year)
data <- list(
            n = n,
            y = y,
            x = x,
            year = year
)
model_string <- textConnection(
  "
  model {
  # Likelihood
  for (i in 1:n){
    y[i] ~ dnorm(mu[i], tau)
    residuals[i] <- y[i] - mu[i]
    mu[i] <- intercept + betas[x[i]] + year[i]*b_year[x[i]]
  }
  # Priors
  intercept ~ dnorm(0, 0.00001)
  betas[1] <- 0
  betas[2] ~ dnorm(0, 0.00001)
  b_year[1] ~ dnorm(0, 0.00001)
  b_year[2] ~ dnorm(0, 0.00001)
  
  sigma ~ dunif(0, 100) # standard deviation
	tau <- 1 / (sigma * sigma) # sigma^2 doesn't work in JAGS
  # Derived values
  residual_sum_sq <- sum(residuals[]^2)
  }
"
)
parameters <- c("intercept","betas","b_year","sigma","residual_sum_sq")
set_initial_value <- function() {
  list( 
    
  )
}
ni <- 10000 ; nt <- 1 ; nb <- 5000 ; nc <- 3
model <- jags(data, set_initial_value, parameters, model_string, n.chains = nc, n.thin = nt, n.iter = ni, n.burnin = nb)
```
### MANOVA
### MANCOVA

## Common Priors

